Reproducible Research: Peer Assessment 1
===========================================

## Loading and preprocessing the data
First we unzip the data. Then we'll load the unziped CSV it into a variable called *data*.  
```{r load_data, echo=TRUE}
unzip("activity.zip")
data <- read.csv("activity.csv")
```
  
Then let's look at the data
```{r examine_data, echo=TRUE}
names(data)
head(data)
str(data)
```

Notice that the date column comes in as a Factor. Let's change that the type to Date. The format is YYYY-MM-DD.

```{r change_date_type, echo=TRUE}
data$date <- as.Date(data$date)
str(data)
```

Great! We should be ready to go!
  


## What is mean total number of steps taken per day?
Let's first get rid of the missing values in the data set.
```{r get_rid_of_missing values, echo=TRUE}
dim(data)
nonmissing <- data[complete.cases(data), ]
dim(nonmissing)
```
We can see that the number of rows in the **nonmissing** dataframe is less than in **data**.

Now let's sum up the number of steps in each day. We can do that with the **aggregate()** method. Then create a histogram showing the distribution of the steps walked in each day.

We can add a bit more context by adding a rug to the histogram.

```{r aggregate_steps_by_date, echo=TRUE}
totals <- aggregate(list(totalStep = nonmissing$steps), by=list(date = nonmissing$date), FUN = sum)
hist(totals$totalStep, breaks = 10, col = "blue", xlab = "Total Steps per day", main = "Distribution of Total Steps per Day")
rug(totals$totalStep)
```

The **mean** is `r toString(mean(totals$totalStep))` steps and the **median** is `r median(totals$totalStep)` steps

## What is the average daily activity pattern?
What we need to do here is aggregate over the interval. So first let's look at what the interval looks like.

```{r explore_interval, echo=TRUE}
unique(nonmissing$interval)
```

Note that the values for the interval identify the five minute interval in day. The format is HHMM, that is to say, 2355 corresponds with 23:55 (11:55 PM.)

So now, let's aggreagate over the interval, to get an average number of steps for each five minute interval (over the days in the data set.) and 

```{r aggregate_over_interval, echo=TRUE}
totalsByInterval <- aggregate(list(avgSteps = nonmissing$steps), by=list(interval = nonmissing$interval), FUN = mean)
head(totalsByInterval)
```

Now let's plot the time series.
``` {r plot_aggr_over_interval, echo=TRUE}
plot(totalsByInterval$avgSteps ~ totalsByInterval$interval, type = "l", col = "red", main = "Average Number of Steps Taken by 5-minute Interval (across all days)", ylab = "Average number of steps", xlab = "Interval (in HHMM format)")
```

Which 5-min interval, on average across all the days in the dataset, contains the maximum number of steps?
``` {r max_number_of_steps, echo=TRUE}
maximum <- totalsByInterval[which.max(totalsByInterval$avgSteps), ]
```
The interval with the max number of steps is `r toString(maximum$interval)`, and the max number of steps is `r toString(maximum$avgSteps)`.

## Imputing missing values
There are many missing values in this data set. Let's take a look at how many observations having missing values.

```{r missing_values, echo=TRUE}
head(data[is.na(data$steps),])
```

There are `r nrow(data[is.na(data$steps),])` rows with missing values. For the first part of this analysis we removed the missing values. However, now we want to impute missing values with that completes the data set, but also doesn't change the distribution of the data. One way we can do this is to replace each observation with missing data with the mean of the steps for the corresponding interval. First what we'll do is copy the dataframe

```{r clone_data, echo=TRUE}
impute <- data.frame(data)
nrow(impute[is.na(impute$steps),])
```

Now we'll complete the missing values in the *impute* dataframe.

```{r impute, echo=TRUE}
library(dplyr)
impute_steps <- function(x) filter(totalsByInterval, interval == x)$avgSteps
impute[is.na(impute$steps),]$steps <- sapply(impute[is.na(impute$steps),]$interval, impute_steps)
missing_rows_imputed <- nrow(impute[is.na(impute$steps),])
```

Note that the number of rows with missing data is now `r toString(missing_rows_imputed)`.

Now we'll aggregate the imputed data set to report the mean and median number of steps per day.

```{r imputed_mean_median, echo=TRUE}
totals_imputed <- aggregate(list(totalSteps = impute$steps), by=list(date = impute$date), FUN = sum)
mean_imputed <- mean(totals_imputed$totalStep)
median_imputed <- median(totals_imputed$totalStep)
```
The **mean** is `r toString(mean_imputed)` steps and the **median** is `r toString(median_imputed)` steps

This is very similar to the totals without the missing values. This is because what we are doing is just replacing a normalized value for the missing values. Furthermore, we are adding the average by interval. In effect, we are not chainging the average of the totals by day.

## Are there differences in activity patterns between weekdays and weekends?
We have a good sense of the activity by date, and by 5-min interval. However, we don't have a good sense of how activity varies from day to day. One might hypothesize that the activity level on a weekend might be more than on a weekday where a subject may be working at relatively sedentary job. In order to get a better sense of the difference between activity on the weekend vs. the weekdays, we need to add a value to our data. We can use the funciton **weekdays()** to help us here.

```{r weeekday_weekends_activity, echo=TRUE}
imputeWithWeekdays <- within(impute, {weekday = ifelse(grepl("S(at|un)", weekdays(impute$date, abbreviate=TRUE)), "weekend", "weekday")})
imputeWithWeekdays$weekday <- as.factor(imputeWithWeekdays$weekday)
```