---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
First we unzip the data. Then we'll load the unziped CSV it into a variable called *data*.  
```{r load_data}
unzip("activity.zip")
data <- read.csv("activity.csv")
```
  
Then let's look at the data
```{r examine_data}
names(data)
head(data)
str(data)
```

Notice that the date column comes in as a Factor. Let's change that the type to Date. The format is YYYY-MM-DD.

```{r change_date_type}
data$date <- as.Date(data$date)
str(data)
```

Great! We should be ready to go!
  


## What is mean total number of steps taken per day?
Let's first get rid of the missing values in the data set.
```{r get_rid_of_missing values}
dim(data)
nonmissing <- data[complete.cases(data), ]
dim(nonmissing)
```
We can see that the number of rows in the **nonmissing** dataframe is less than in **data**.

Now let's sum up the number of steps in each day. We can do that with the **aggregate()** method. Then create a histogram showing the distribution of the steps walked in each day.

We can add a bit more context by adding a rug to the histogram.

```{r aggregate_steps_by_date}
totals <- aggregate(list(TotalStep = nonmissing$steps), by=list(Date = nonmissing$date), FUN = sum)
hist(totals$TotalStep, breaks = 10, col = "blue", xlab = "Total Steps per day", main = "Distribution of Total Steps per Day")
rug(totals$TotalStep)
```

The **mean** is `r toString(mean(totals$TotalStep))` steps and the **median** is `r median(totals$TotalStep)` steps

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
